Incremental Forward Reasoning for White-Box~Proof~Search

### 2025-02-14 paper.tex
Edit example 2

\paragraph{Example 2.}
We now consider another example with more than one shared variable.
Let $T_1, T_2$ be two proposition that both depend on the variable $x$ and $y$.
For simplicity, the rule considered will be $T_1(x,y) \to T_2(x,y) \to R$.
The hypotheses added to the context will be, in order, $h_1 : T_2(a,a)$, $h_2 : T_1(a,b)$ and $h_3 : T_2(a,b)$.
First, $h_1 : T_2(a,a)$ is considered :

% NOTE : remove comments to get full trees.
\begin{tikzpicture}[level distance=12mm, scale=0.50]
    \node[xshift = 30] {$\eta_x$}
        %child[xshift = -20] {node {1}
        %  child {node {$a$}
        %    child {node {$\{ \}$}}}
        %  child {node {$b$}
        %    child {node {$\{ \}$}}}}
        child[xshift = 40] {node {2}
        child {node {$a$}
            child {node {$\{h_1\}$}}}
        % child {node {$b$}
        %    child {node {$\{ \}$}}}
        };
    \node[xshift = 110] {$\eta_y$}
        %child[xshift = -20] {node {1}
        %    child {node {$a$}
        %        child {node {$\{ \}$}}}
        %    child {node {$b$}
        %        child {node {$\{ \}$}}}}
        child[xshift = 40] {node {2}
            child {node {$a$}
                child {node {$\{h_1\}$}}}
        %    child {node {$b$}
        %        child {node {$\{ \}$}}}
        };
    \node[xshift = 190, yshift = -25] {$\mu_x$};
        %child[xshift = -20] {node {1}
        %    child {node {$a$}
        %        child {node {$\{ \}$}}}
        %    child {node {$b$}
        %        child {node {$\{ \}$}}}}
        %child[xshift = 20] {node {2}
        %    child {node {$a$}
        %        child {node {$\{\}$}}}
        %    child {node {$b$}
        %        child {node {$\{ \}$}}}
        %};
    \node[xshift = 270, yshift = -25] {$\mu_y$};
        %child[xshift = -20] {node {1}
        %    child {node {$a$}
        %        child {node {$\{ \}$}}}
        %    child {node {$b$}
        %        child {node {$\{ \}$}}}}
        %child[xshift = 20] {node {2}
        %    child {node {$a$}
        %        child {node {$\{\}$}}}
        %    child {node {$b$}
        %        child {node {$\{ \}$}}}
        %};
\end{tikzpicture}

At this point, there are no matches so there is nothing to extend.
We add the next hypothesis in the queue, that is $h_2 : T_1(a,b)$, to the variable maps :

% NOTE : uncomment to get full trees.
\begin{tikzpicture}[outer sep=auto, level distance=12mm, scale=0.50]
    \node[xshift = 0] {$\eta_x$}
        child[xshift = -20] {node {1}
            child {node {$a$}
                child {node {$\{h_2\}$}}}
            %child {node {$b$}
            %    child {node {$\{ \}$}}}
            }
        child[xshift = 20] {node {2}
            child {node {$a$}
                child {node {$\{h_1\}$}}}
            %child {node {$b$}
            %    child {node {$\{ \}$}}}
            };
    \node[xshift = 80] {$\eta_y$}
        child[xshift = -20] {node {1}
            %child {node {$a$}
            %    child {node {$\{ \}$}}}
            child {node {$b$}
                child {node {$\{ h_2\}$}}}}
        child[xshift = 20] {node {2}
            child {node {$a$}
                child {node {$\{h_1\}$}}}
            %child {node {$b$}
            %    child {node {$\{ \}$}}}
                };
    \node[xshift = 160, yshift = 0] {$\mu_x$}
        child[xshift = -20] {node {1}
            child {node {$a$}
                child {node {$\{[h_2]\}$}}}
            %child {node {$b$}
            %   child {node {$\{ \}$}}}
                };
        %child[xshift = 20] {node {2}
        %    child {node {$a$}
        %        child {node {$\{\}$}}}
        %    child {node {$b$}
        %        child {node {$\{ \}$}}}
        %};
    \node[xshift = 240, yshift = 0] {$\mu_y$}
        child[xshift = -20] {node {1}
        %    child {node {$a$}
        %        child {node {$\{ \}$}}}
            child {node {$b$}
                child {node {$\{[h_2]\}$}}}
        };
        %child[xshift = 20] {node {2}
        %    child {node {$a$}
        %        child {node {$\{\}$}}}
        %    child {node {$b$}
        %        child {node {$\{ \}$}}}
        %};
\end{tikzpicture}

As this hypothesis is in the first slot, we create the corresponding level 1 partial match $m_0 = {(h_2 : T_1(x, y))}$ for which we check for suitable hypotheses that could extend the match.
Note that  $s(lvl(m_0) + 1) = s(2)$ which is the intersection of the variables appearing in the slots $i < 2$ with the variables appearing in slot $2$.
Thus $s(2) = \{x,y\}$ and we compute:
\[
  \bigcap_{x\in s(lvl(m_0) + 1)} \eta_x \left(i + 1, \sigma_{m_0}(x)\right) = \eta_x(2, a) \cap \eta_y(2, b) = \emptyset
\]
Since the set is empty, the match can not be currently extended.
We turn to the remaining hypothesis $h_3 : T_2(a,b)$ in the queue and add it to the maps:

\begin{tikzpicture}[outer sep=auto, level distance=12mm, scale=0.50]
\node[xshift = 0] {$\eta_x$}
    child[xshift = -20] {node {1}
        child {node {$a$}
            child {node {$\{h_2\}$}}}
        %child {node {$b$}
        %    child {node {$\{ \}$}}}
        }
    child[xshift = 20] {node {2}
        child {node {$a$}
            child {node {$\{h_1, h_3\}$}}}
        %child {node {$b$}
        %    child {node {$\{ \}$}}}
        };
\node[xshift = 80] {$\eta_y$}
    child[xshift = -20] {node {1}
        %child {node {$a$}
        %    child {node {$\{ \}$}}}
        child {node {$b$}
            child {node {$\{h_2\}$}}}}
    child[xshift = 20] {node {2}
        child {node {$a$}
            child {node {$\{h_1\}$}}}
        child {node {$b$}
            child {node {$\{h_3\}$}}}};
    \node[xshift = 160, yshift = 0] {$\mu_x$}
        child[xshift = -20] {node {1}
            child {node {$a$}
                child {node {$\{[h_2]\}$}}}
            %child {node {$b$}
            %   child {node {$\{ \}$}}}
                }
        child[xshift = 20] {node {2}
            child {node {$a$}
                child {node {$\{[h_2,h_3]\}$}}}
        %    child {node {$b$}
        %        child {node {$\{ \}$}}}
        };
    \node[xshift = 240, yshift = 0] {$\mu_y$}
        child[xshift = -20] {node {1}
        %    child {node {$a$}
        %        child {node {$\{ \}$}}}
            child {node {$b$}
                child {node {$\{[h_2]\}$}}}
        }
        child[xshift = 20] {node {2}
        %    child {node {$a$}
        %        child {node {$\{\}$}}}
            child {node {$b$}
                child {node {$\{[h_2,h_3]\}$}}}
        };
\end{tikzpicture}

We are now ready to look for potential matches that could be extended by $h_3$.
In this particular case, since $h_3$ is in slot $2$ we look for $s(2) = \{x,y\}$.
Then,
\[
\bigcap_{x\in s(2)} \mu_x \left(i - 1, \sigma_h(x)\right) = \mu_x(1, a) \cap \mu_y(1, b) = \{m_0\}
\]

We have found two compatible hypotheses for our rule and we can then add the result $R$ as hypothesis to the context.

>>>>

\paragraph{Example 2.}
We now consider a rule $s : \All{x~y}{A~x~y → B~x~y → C~x~y}$ with multiple shared variables.
We add hypotheses $h₁ : B~a~a$, $h₂ : A~a~b$ and $h₃ : B~a~b$ to the rule state of $s$, starting with $h₁$:
% NOTE : remove comments to get full trees.
\begin{center}
  \begin{tikzpicture}[level distance=12mm, scale=0.50]
    \node[xshift = 30] {$H_x$}
        %child[xshift = -20] {node {1}
        %  child {node {$a$}
        %    child {node {$\{ \}$}}}
        %  child {node {$b$}
        %    child {node {$\{ \}$}}}}
        child {node {2}
        child {node {$a$}
            child {node {$\{h_1\}$}}}
        % child {node {$b$}
        %    child {node {$\{ \}$}}}
        };
    \node[xshift = 110] {$H_y$}
        %child[xshift = -20] {node {1}
        %    child {node {$a$}
        %        child {node {$\{ \}$}}}
        %    child {node {$b$}
        %        child {node {$\{ \}$}}}}
        child {node {2}
            child {node {$a$}
                child {node {$\{h_1\}$}}}
        %    child {node {$b$}
        %        child {node {$\{ \}$}}}
        };
    \node[xshift = 190] {$M_x$};
        %child[xshift = -20] {node {1}
        %    child {node {$a$}
        %        child {node {$\{ \}$}}}
        %    child {node {$b$}
        %        child {node {$\{ \}$}}}}
        %child[xshift = 20] {node {2}
        %    child {node {$a$}
        %        child {node {$\{\}$}}}
        %    child {node {$b$}
        %        child {node {$\{ \}$}}}
        %};
    \node[xshift = 270] {$M_y$};
        %child[xshift = -20] {node {1}
        %    child {node {$a$}
        %        child {node {$\{ \}$}}}
        %    child {node {$b$}
        %        child {node {$\{ \}$}}}}
        %child[xshift = 20] {node {2}
        %    child {node {$a$}
        %        child {node {$\{\}$}}}
        %    child {node {$b$}
        %        child {node {$\{ \}$}}}
        %};
  \end{tikzpicture}
\end{center}
At this point there are no matches, so there is nothing to extend.
We now add $h_2 : A~a~b$, which initially generates a match $m = \{1 ↦ h₂\}$ that is added to $M$:
% NOTE : uncomment to get full trees.
\begin{center}
  \begin{tikzpicture}[outer sep=auto, level distance=12mm, scale=0.50]
    \node[xshift = 0] {$H_x$}
        child {node {2}
            child {node {$a$}
                child {node {$\{h_1\}$}}}
            %child {node {$b$}
            %    child {node {$\{ \}$}}}
            };
    \node[xshift = 80] {$H_y$}
        child {node {2}
            child {node {$a$}
                child {node {$\{h_1\}$}}}
            %child {node {$b$}
            %    child {node {$\{ \}$}}}
                };
    \node[xshift = 160, yshift = 0] {$M_x$}
        child {node {1}
            child {node {$a$}
                child {node {$\{[h_2]\}$}}}
            %child {node {$b$}
            %   child {node {$\{ \}$}}}
                };
        %child[xshift = 20] {node {2}
        %    child {node {$a$}
        %        child {node {$\{\}$}}}
        %    child {node {$b$}
        %        child {node {$\{ \}$}}}
        %};
    \node[xshift = 240, yshift = 0] {$M_y$}
        child {node {1}
        %    child {node {$a$}
        %        child {node {$\{ \}$}}}
            child {node {$b$}
                child {node {$\{[h_2]\}$}}}
        };
        %child[xshift = 20] {node {2}
        %    child {node {$a$}
        %        child {node {$\{\}$}}}
        %    child {node {$b$}
        %        child {node {$\{ \}$}}}
        %};
  \end{tikzpicture}
\end{center}
To find hypotheses compatible with $m$, we consider the set of variables $\vars(\lvl(m) + 1)$.
These are the variables shared between slot $\lvl(m) + 1 = 2$ and all previous slots, i.e.\ $x$ and $y$.
Hence the compatible hypotheses are
\[
  \bigcap_{x \in \vars(\lvl(m) + 1)} H(x, i + 1, \sub(m)(x)) = H(x, 2, a) \cap H(y, 2, b) = \emptyset
\]
This means the match cannot currently be extended.
We now turn to the remaining hypothesis $h_3 : B~a~b$, adding it first to $H$:
\begin{center}
  \begin{tikzpicture}[outer sep=auto, level distance=12mm, scale=0.50]
    \node[xshift = 0] {$H_x$}
      child {node {2}
          child {node {$a$}
              child {node {$\{h_1, h_3\}$}}}
          %child {node {$b$}
          %    child {node {$\{ \}$}}}
          };
    \node[xshift = 80] {$H_y$}
      child {node {2}
          child {node {$a$}
              child {node {$\{h_1\}$}}}
          child {node {$b$}
              child {node {$\{h_3\}$}}}};
    \node[xshift = 160, yshift = 0] {$M_x$}
      child {node {1}
          child {node {$a$}
              child {node {$\{[h_2]\}$}}}
          %child {node {$b$}
          %   child {node {$\{ \}$}}}
              };
    \node[xshift = 240, yshift = 0] {$M_y$}
      child {node {1}
      %    child {node {$a$}
      %        child {node {$\{ \}$}}}
          child {node {$b$}
              child {node {$\{[h_2]\}$}}}
      };
  \end{tikzpicture}
\end{center}
Since $h₃$ matches slot $i = 2$, the matches that can be extended by $h₃$ are
\[
  \bigcap_{x \in \vars(i)} M(x, i - 1, \sub_{i}(x)) = M(x, 1, a) \cap M(y, 1, b) = \{[h₂]\}
\]
We have found the complete match $[h₂, h₃]$ and can add a new hypothesis of type $C~a~b$.
